\pkg{DNest4}: Diffusive Nested Sampling in \proglang{C++} and \proglang{Python}

### 2016-05-31 paper/ms.tex
Direct method for log-uniform prior, and why it's bad

In the example, the proposal for $\sigma$ is implemented by taking advantage
of the uniform prior for $\ln(\sigma)$. So $\sigma$ is transformed by
taking a logarithm, a proposal move is made (that satisfies detailed balance
with respect to a uniform distribution), and then $\sigma$ is exponentiated
again. The step for the uniform prior between -10 and +10
uses the following code:
\begin{CodeChunk}
\begin{CodeInput}
sigma += 20.0*rng.randh();
DNest4::wrap(sigma, -10.0, 10.0);
\end{CodeInput}
\end{CodeChunk}
The factor of 20 accounts for the prior width, and
the \code{wrap(double\&, double, double)} function uses the modulo operator
to construct periodic boundaries. For example, if the perturbation results
in a value of 10.2, which is outside the prior range, the value is modified to
$-9.8$.
The \code{wrap} function has no return value and works by modifying its
first argument, which is passed by reference.

It is imperative that \code{from_prior} and
\code{perturb}
be consistent with each other. One technique
for verifying this is to sample the prior for a long time
(by setting the maximum number of levels to 1) and inspect
{\tt sample.txt} to ensure that each parameter is exploring
the prior correctly.

>>>>

In the example, the proposal for $\sigma$ is implemented by taking advantage
of the uniform prior for $\ln(\sigma)$. So $\sigma$ is transformed by
taking a logarithm, a proposal move is made (that satisfies detailed balance
with respect to a uniform distribution), and then $\sigma$ is exponentiated
again. The step for the uniform prior between -10 and +10
uses the following code:
\begin{CodeChunk}
\begin{CodeInput}
sigma += 20.0*rng.randh();
DNest4::wrap(sigma, -10.0, 10.0);
\end{CodeInput}
\end{CodeChunk}
The factor of 20 accounts for the prior width, and
the \code{wrap(double\&, double, double)} function uses the modulo operator
to construct periodic boundaries. For example, if the perturbation results
in a value of 10.2, which is outside the prior range, the value is modified to
$-9.8$.
The \code{wrap} function has no return value and works by modifying its
first argument, which is passed by reference.
Alternatively, the log-uniform prior, which has density proportional
to $1/\sigma$, could have been used directly by not using the log/exponential
trick and adding $\ln\left[(1/\sigma')/(1/\sigma)\right]$ to the
return value. However, this isn't recommended for a distribution like this
which covers several orders of magnitude, because the appropriate scale size
for the proposal is less clear. This is likely to cause inefficient
sampling.

\subsection{Consistency of prior and proposal}
It is imperative that \code{from_prior} and
\code{perturb}
be consistent with each other, and that each implements
the prior distributions that you want to use. One technique
for testing this is to sample the prior for a long time
(by setting the maximum number of levels to 1) and inspect
{\tt sample.txt} to ensure that each parameter is exploring
the prior correctly.
